---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# sudoku

The goal of sudoku is to provide simple functions for solving sudoku puzzles in R.

## Installation

You can install sudoku from github with:

``` r
# install.packages("devtools")
devtools::install_github("alexpavlakis/sudoku")
```


## Example

The workhorse function is `solve_sudoku`, which takes two arguments:

* `sudoku_matrix`: an unsolved sudoku puzzle in matrix form, with `NA` in unknown cells;
* `verbose`: `TRUE` if you want the function to print intermediate steps (defaults to `FALSE`)

`check_integrity` will tell you if a puzzle is completed correctly.

solve sudoku first attempts to solve the sudoku with basic sudoku logic.  If this does not work, it uses a backtracking algorithm to find a solution (if one exists).  Core functions (such as backtracking) are written in C++ for speed.

```{r, puzzle}
# An unsolved puzzle
print(sudoku)

# Solve
solved_puzzle <- solve_sudoku(sudoku)
print(solved_puzzle)

# Check that it is correct
check_integrity(solved_puzzle)
```

`solve_sudoku` completes puzzles in the blink of an eye.  The example below shows the time taken to solve an easy puzzle (49 empty cells) and a hard puzzle (59 empty cells).  The easy puzzle is solved in 10-20 milliseconds and the hard puzzle is solved in 150-200 milliseconds.

```{r showpuzzles}
# Easy puzzle - 49 empty cells
sum(is.na(sudoku))
print(sudoku)

# Hard Puzzle - 59 empty cells
sum(is.na(hard_sudoku))
print(hard_sudoku)

```
```{r rmb, warning=F, message=F, results=F}
library(microbenchmark)
m <- microbenchmark(easy = solve_sudoku(sudoku),
                    hard = solve_sudoku(hard_sudoku), 
                    times = 100, unit = 's')
```
```{r showmb}
print(m, digits = 3)
```
