---
output: github_document
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/alexpavlakis/sudokuplyr.svg?branch=master)](https://travis-ci.org/alexpavlakis/sudokuplyr)
<!-- badges: end -->

# sudokuplyr

The goal of sudokuplyr is to provide simple functions for solving, creating, and analyzing sudoku puzzles in R.



## Installation

You can install sudokuplyr from github with:

``` r
# install.packages("devtools")
devtools::install_github("alexpavlakis/sudokuplyr")
```

## Examples

The following methods are available for objects of class `sudoku`:

* `print`
* `plot`
* `solve`
* `summary`


```{r example, fig.height=6, fig.width=6, fig.align='center'}
# Object of class sudoku
class(sudoku)

# Print
print(sudoku)

# Solve
solve(sudoku)

# Plot
plot(sudoku)

# Summarize
summary(sudoku)
```

`solve_sudoku`, `print_sudoku`, `analyze_sudoku`, and `plot_sudoku` operate on matrices directly.

`solve_sudoku` completes puzzles in the blink of an eye.  The example below shows the time taken to solve an easy puzzle (49 empty cells) and a hard puzzle (59 empty cells).  Both are solved in a fraction of a millisecond on a MacBook Air.

```{r rmb, warning=F, message=F, results=F}

m <- microbenchmark(easy = solve(sudoku),
                    hard = solve(hard_sudoku), times = 100, unit = 'ms')
```
```{r showmb}
print(m, digits = 2)
```

`generate_sudoku` creates randomly generated complete sudoku puzzles.  The `seed` argument can be used to create reproducible random puzzles or left `NULL` (default).  `generate_puzzle` creates randomly generated incomplete sudoku puzzles with a specified number of clues.

```{r gensudoku}
new_puzzle <- generate_puzzle(clues = 32, unique = TRUE, seed = 56)
print(new_puzzle)
```


`get_all_solutions` generates all possible solutions for a given puzzle.  Puzzles with fewer than 20-25 clues (depending on the puzzle, of course), can have _a lot_ of solutions, and this function can take a while to find them all.  Beware.  Most puzzles that appear in newspapers and magazines have only one solution.

```{r generate_all}
# This puzzle has a lot of solutions
puzzle <- generate_puzzle(clues = 28, unique = FALSE, seed = 56)
all_solutions <- get_all_solutions(puzzle, stop_early = FALSE)
length(all_solutions)
```

`
