---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# sudoku

The goal of sudoku is to provide simple functions for solving most sudoku puzzles.

## Installation

You can install sudoku from github with:

``` r
# install.packages("devtools")
devtools::install_github("alexpavlakis/sudoku")
```


## Example

The workhorse function is `solve_sudoku`, which takes two arguments:

* `sudoku_matrix`: an unsolved sudoku puzzle in matrix form, with `NA` in unknown cells;
* `verbose`: `TRUE` if you want the function to print intermediate steps (defaults to `FALSE`)

`check_integrity` will tell you if a puzzle is completed correctly.

solve sudoku first attempts to solve the sudoku with basic sudoku logic.  If this does not work, it uses a backtracking algorithm to find a solution (if one exists).

```{r, puzzle}
# An unsolved puzzle
print(sudoku)

# Solve
solved_puzzle <- solve_sudoku(sudoku)

# Check that it is correct
check_integrity(solved_puzzle)
```
